Requantization of time-dependent mean field for pairing collective motion in superfluid nuclei

### 2019-01-09 thesis/thesis.tex
thesis

Of course, single-particle motion picture cannot describe the nuclear system perfectly. To describe the nuclear structure more precisely, residual interactions need to be taken account. In 1950s, nuclear physicists found one of the most important residual interactions, pairing correlation. One of the clear evidences for pairing correlation detected from experiment is the separation energy difference between odd nuclei and even nuclei. Fig. \ref{S_n} shows the separation energies for Sn, Sb, and Te isotopes as a function of neutron number. The isotopes in even number of neutron have about $2\sim3$ MeV larger separation energies than the isotopes in odd number of neutron.
It is odd-even staggering (OES) effect, which is detected in whole mass region of nuclei (Fig. \ref{gap}). In this figure, $2\Delta$ corresponds to the separation energy difference between even-even nuclei and odd mass number nuclei. The approximate value is $2\Delta \approx 24A^{-1/2}$ MeV. It indicates a two-body attractive interaction should be included based on single-particle motion picture. In addition, if we examine the spin-parity of the ground states, we can find all of the even-even nuclei are $J^{\pi}=0^+$ states and most of the odd mass number nuclei have the same spin-parity as their single-particle orbit which the last nucleon fulfills in.
To explain these phenomena, it is natural to consider that the ground state of even-even nuclei consist of coupled $J^{\pi}=0^+$ pairs and the ground state of odd mass number nuclei consist of coupled $J^{\pi}=0^+$ pairs except one unpaired nucleon in last single-particle orbit. The two-body attractive interaction which couples two identical nucleons into $J^{\pi}=0^+$ state is called pairing correlation.

@@

First, we discuss the pairing rotation. In classical picture, pairing rotation is the motion which the system rotates in the U(1) gauge space (red bi-direction arrow in Fig. \ref{pair_rot}). The motion is described by the coordinate, gauge angle $\Phi$, and the corresponding momentum, total particle number $N$. 
%It is similar to normal rotation in 3D space which described by the angles ($\theta_x, \theta_y, \theta_z$) and their conjugate angular momenta ($J_x, J_y, J_z$). 
Of course, $N$ is a constant of motion. In quantum picture, if we assume the ground state of $N_0$ particle system as the ``ground state'', the ground states in $N=N_0\pm2, N_0\pm4,\cdots$ correspond to the ``excited states'' in the pairing rotational mode. The excitation energy is $E_{\rm rot}^{\rm pair}=\frac{(N-N_0)^2}{2\mathcal{J}}$, where $\mathcal{J}$ is the pairing-rotational moment of inertia. In experiment, the pairing rotational band is expected to be observed in even-even isotope chain or even-even isotone chain in open-shell nuclei. A clear observation is found in Sn isotopes shown in Fig. \ref{pair_rot}. Except a constant value and linear term with respect to mass number $A$, we can find the energies of ground states consist of a parabolic line - pairing rotational band. Recently, the systematic study for the pairing-rotational moment of inertia has been done in Ref. \cite{HN16}. The experimental moments of inertia obtained from two neutron(proton) separation energies have excellent correspondence with the theoretical moments of inertia obtained from NG mode in quasiparticle random phase approximation (QRPA) \cite{TV61}, not only for the spherical nuclei, but also for the deformed nuclei. Because the pairing-rotational moments of inertia are free from ambiguities caused from odd-even system, they are supposed to be excellent indicators of pairing correlation rather than pairing gaps.

@@

The signatures for pairing collective motion can also be observed from the two-neutron(proton) transfer reaction \cite{BHR73}, such as $(p, t)$, $(t, p)$, and $({}^3{\rm He}, n)$. The cross sections of two particle transfer reaction strongly depend on the matrix elements of $J^{\pi}=0^+$ pair-transition operators (e.g. pair-additional transition: $\sigma( P_{ad};0_{\alpha}^+(N) \to 0_{\beta}^+(N+2))\sim |\braket{N+2,\beta|\sum_k a^{\dag}_k a^{\dag}_{\bar{k}}|N,\alpha}|^2$). From theoretical prediction, the pairing correlation is known to enhance $\sigma(g.s.\to g.s.)$ and decrease $\sigma(g.s.\to 0_{ex}^+)$ significantly, hence, $\sigma(g.s.\to 0_{ex}^+)/\sigma(g.s.\to g.s.)\ll 1$ is expected in superfluid system. One example of $(p, t)$ reactions is demonstrated in Fig. \ref{pair_rot}. As mentioned in previous paragraph, the ground states of even-even Sn isotopes consist of a pairing rotational band. In such situation, the intraband transition is dominant with respect to the interband transition. Therefore, the cross sections of two-neutron transfer reaction is consistent with theoretical prediction. In general, the ratio $\sigma(g.s.\to 0_{ex}^+)/\sigma(g.s.\to g.s.)$ for two-neutron(proton) transfer reaction are excellent indicators to describe the collectivity of pairing dynamics.

@@

\begin{figure}[tb]
 \begin{center}
    \includegraphics[width=100mm]{images/phonon.eps}
 \end{center}
 \begin{center}
    \includegraphics[width=100mm]{images/112Cd.eps}
 \end{center}
  \caption{$B(E2)$ values, relative to the $B(E2; 2_1^+\to 0_1^+)$ value, and energy spectra in phonon excitation picture. Upper panel: expected $B(E2)$ values for a quadrupole harmonic vibrator. Lower panel: data observed for ${}^{112}$Cd. Dashed transitions indicate an unobserved transition. Detailed information is discussed in Ref. \cite{GW10}.}
  \label{112Cd}
\end{figure}
The mystery of excited $0^+$ states exist not only in deformed nuclei, but also in spherical nuclei. A best example is the structure of $0_2^+$ or $0_3^+$ states in Cd isotopes observed in recent experiment \cite{GW10}. The lower part of Fig. \ref{112Cd} shows the excitation energy and $B(E2)$ transition strength in ${}^{112}$Cd. Comparing with the typical values for the ideal quadrupole harmonic vibration in spherical nuclei shown in the upper part of Fig. \ref{112Cd}, we can find that the observed excitation energy spectra have good coincidence with phonon excitation picture. The $0_2^+$ or $0_3^+$ states seem to be interpreted as two-phonon or three-phonon excitation. However, the $B(E2)$ values are quite different between the two, especially for the no observed decay from $0_{2ph}^+$ to $2_1^+$ state. Such phenomena are also seen in other Cd isotopes. The configuration of these $0^+$ states are supposed to be not pure phonon excitation, but proton pair-excitation between $2p_{1/2}$ and $1g_{9/2}$ orbits from the hint of the ground state configurations of Ag isotopes \cite{HW11}. The pairing should play an important role if the excited $0^+$ states are described by pair-excitation.

@@

%Large amplitude motion, Adiabatic approach
To describe the low-lying excited states, especially for the excited $0^+$ states, we need to introduce the large-amplitude motions beyond the linear regime.
Collective model, which is designed by a couple of collective coordinates and their collective potential and mass parameters, is useful for describing the large-amplitude dynamics. The most popular phenomenological collective model is Bohr collective model, five-dimensional collective Hamiltonian (5DCH) with the quadrupole deformation parameters $(\beta,\gamma)$ and the Euler angles $(\phi,\theta,\psi)$. It has been applied to describe the low-lying excited states in realistic nuclei, with the construction of the collective potential and mass parameters from the microscopic degrees of freedom \cite{NMMY16}. However, the pairing degrees of freedom are not explicitly included in the Bohr collective model. As discussed in Sec. \ref{1-1-2}, the pairing vibrational degrees of freedom supposed to be a key ingredient to understand the low-lying  excited $0^+$ states. We have great interest for the collective coordinates describing the pairing corrective motions.

@@

\clearpage{\pagestyle{empty}\cleardoublepage}
\chapter{Requantization of TDHFB in integrable system}
\label{integrable}
As discussed in Sec. \ref{constant}, there is only one time-dependent degree of freedom in two-level system. We know that one-dimensional system must be integrable system.
We obtained two achievements from the study for the integrable system. First, the requantization method, stationary-phase approximation to the path integral, can be applied straightforward to describe the collective excited $0^+$ states. We find that SPA is superior to the other requantization methods, especially for the calculation of two-particle transition strength. Second, the collective coordinates are explicit in integrable system. Using the explicit collective coordinates, we can examine the performance of the collective model which assumes the pairing gap parameter as the collective coordinate. The conclusion is the pairing gap parameter can only describe a part of collective space, and the applicability of the collective model is limited.

@@

We show an example of the classical trajectories in two-level system. Fig. \ref{fig:phase_space} is the contour lines fulfilling (\ref{contour}), which correspond to the classical trajectories, with $\Omega_1=\Omega_2=\Omega=8$, $\nu=0$, and $N=2\Omega=16$ with various coupling constant $g$. The dimensionless coupling constant $x$ is defined as $x=2g\Omega/(\epsilon_2-\epsilon_1)$ (See Appendix \ref{two-level}). The phase transition from normal state to superfluid state occurs at $x=1$ in closed shell ($N=2\Omega$). At $x<1$ corresponding the left panel in Fig. \ref{fig:phase_space}, the trajectory of ground state is $j=-4$ independent from $\phi$. It is consistent with the fully occupied lower level $n_1=N$ and the empty upper level $n_2=0$. For the dynamical states, the trajectories are open with gentle slopes. At $x>1$ corresponding the middle and right panels in Fig. \ref{fig:phase_space}, the trajectory of ground state becomes a point at $\phi=0$, $j=j_0$ $(-4<j_0<4)$. With superfluid ground state, the dynamical states have two types, closed trajectory and open trajectory. The closed trajectory appears around $(\phi,j)=(0,j_0)$ at low-energy region. On the other hand, the open trajectory with steep slope appears at high-energy region relatively. The stronger pairing correlation, the higher excitation energy of the transition point from closed trajectory to open trajectory. The structure of the excited states between closed trajectory and open trajectory is quite different. They influence the matrix element of two-particle transfer reaction significantly. We will discuss the feature of the excited states in the following sections.

\subsection{TDHFB dynamics in adiabatic\footnote{In this subsection, the word ``adiabatic'' means that the collective velocity is small.} limit}
\label{ATDHFB}
In general collective motion, the frequency of the collective motion is much smaller than the frequency in single particle motion ($\hbar\omega\ll\epsilon_m-\epsilon_i$). The slow velocity of the collective motion enable us to expand the TDHFB Hamiltonian up to second order with respect to the collective momentum. We discuss the behavior of TDHFB dynamics in adiabatic approximation.\par
From (\ref{Hamiltonian2}), we find that the classical Hamiltonian is even function with respect to $\phi$. With time reversal $t\to-t$, (\ref{TDHFB_equation}) is invariant under $j\to j, \phi\to -\phi$, which indicates that $j$ is time-even and $\phi$ is time-odd. Thus, we switch the coordinate and momentum 
\begin{equation}
	(\phi,j)\to(j,-\phi) ,
	\label{switch}
\end{equation}
to make the coordinate time-even. The classical Hamiltonian under adiabatic approximation becomes
\begin{align}
  \mathcal{H}(j,\phi) &\approx V(j) + \frac{1}{2}B^{-1}(j)\phi^2,
  \label{ATDHFB_Hamiltonian}
\end{align}
where 
\begin{align}
	V(j) = \mathcal{H}(\phi=0,j) 
	=& \sum_{\alpha=1,2} \epsilon_{\alpha}\nu_{\alpha} + \sum_{\alpha=1,2} 2\epsilon_{\alpha}j_{\alpha} - g\sum_{\alpha=1,2} \left( (\Omega_{\alpha}-\nu_{\alpha}) j_{\alpha} - j_{\alpha}^2 +\frac{j_{\alpha}^2}{\Omega_{\alpha}-\nu_{\alpha}} \right) \nonumber 
	\label{potential_two_level} \\
	&- 2g\sqrt{j_1j_2(\Omega_{1}-\nu_{1}-j_{1})(\Omega_{2}-\nu_{2}-j_{2})} \\
	B^{-1}(j) = \left. \frac{\partial^2\mathcal{H}}{\partial\phi^2} \right|_{\phi=0}
	=& 2g\sqrt{j_1j_2(\Omega_{1}-\nu_{1}-j_{1})(\Omega_{2}-\nu_{2}-j_{2})} .
	\label{mass_two_level}
\end{align}
It is obvious that $V(j)$ is potential and $B(j)$ is mass parameter. We define a classical trajectory obtained from (\ref{ATDHFB_Hamiltonian}) as adiabatic TDHFB (ATDHFB) trajectory. Because the closed trajectories appear around the energy minimum point $(\phi,j)=(0,j_0)$, we expect that adiabatic approximation performs excellent for closed trajectories.
\begin{figure}[tb]
 \begin{center}
  \includegraphics[width=60mm,angle=-90]{images/phase_compare.eps}
 \end{center}
 \caption{Energy contour plot in the same system as Fig. \ref{fig:phase_space}. Black solid lines correspond to TDHFB trajectories and gray dashed lines correspond to ATDHFB trajectories. 
	}
 \label{fig:phase_compare}
\end{figure}

We compare the difference of the classical trajectories between TDHFB and ATDHFB. Fig. \ref{fig:phase_compare} shows both TDHFB and ATDHFB trajectories in the same system as in Fig. \ref{fig:phase_space}. For weak pairing correlation ($x<1$), both results are similar, especially at the low-energy region. For strong pairing correlation ($x>1$), both results are almost identical for closed trajectories. The opened trajectories are also well reproduced near the closed trajectories. The deviation becomes large at the high-energy region. In conclusion, ATDHFB is excellent approximation at low-energy region for both open and closed trajectories.

\section{Requantization methods}
\label{3-2}
To obtain quantum states from TDHFB dynamics, quantization of the TDHFB trajectories, namely, requantization is necessary procedure. We emphasize that there are infinite quantum pictures corresponding to one classical picture. Different quantization methods lead to different quantum states. The straightforward requantization method is canonical quantization, which is commonly applied into nuclear collective model. Here, we apply the canonical quantization to describe the excited states. However, we find that canonical quantization has serious problems to describe the pairing dynamics. Such problems are discussed in Sec. \ref{3-3}. To solve or avoid the problems, alternative requantization methods, Fourier decomposition and stationary-phase approximation to the path integral, are also introduced. In this section, we explain each requantization method in detail, respectively.

@@

Before ending this section, let us note the periodic conditions
of the coordinates and the quantization condition.
Since the two original variables, $(\chi_1,\chi_2)$,
are independent periodic variables of the period $2\pi$,
in addition to the trivial periodicity of $2\pi$ for $\Phi$ and of $4\pi$
for $\phi$,
we have periodic conditions for
$(\Phi,\phi)\rightarrow (\Phi\pm \pi,\phi\pm 2\pi)$ and
$(\Phi,\phi)\rightarrow (\Phi\pm \pi,\phi\mp 2\pi)$.
The former (latter) corresponds to $\chi_2\pm 2\pi$ ($\chi_1\pm 2\pi$)
with $\chi_1$ ($\chi_2$) being fixed.
For open TDHFB trajectory (e.g., Fig.~\ref{fig:phase_space}),
the quantization condition becomes
\begin{align}
 \mathcal{T}_{J,E_k}(\Phi\pm\pi,\phi\pm 2\pi;J) &= \mathcal{T}_{J,E_k}(\Phi,\phi;J) + 2m\pi \nonumber \\
	 \Leftrightarrow \quad\quad \pm J \pi + \int_{-\pi}^\pi j d\phi
	&= 2m\pi .
\end{align}
This leads to the following:
\begin{equation}
	\int_{-\pi}^\pi j d\phi = 
	\begin{cases}
		2k\pi & \mbox{ for } J=2n ,\\
		(2k+1)\pi & \mbox{ for } J=2n+1 ,
	\end{cases}
	\label{EBK_2}
\end{equation}
where $m$, $k$ and $n>0$ are integer numbers.

@@

From the study of various requantization methods for the two-level pairing model,
canonical quantization and Fourier decomposition often fail to produce an approximate value to the exact solution.
In contrast,
the stationary-phase approximation (SPA) to the path integral \cite{SM88} 
can give quantitative results not only for the excitation energies, 
but also for the wave functions and two-particle-transfer strengths.
The quantized states obtained in the SPA have two advantages;
first, the wave functions are given directly in terms of
the microscopic degrees of freedom, and second,
the restoration of the broken symmetries are automatic.
In the pairing model, the quantized states are eigenstates of the particle-number
operator.
On the other hand, applications of the SPA have been limited to
integrable systems.
This is because we need to find separable periodic trajectories on a classical
torus.
Since the nuclear systems, of course, correspond to non-integrable systems,
a straightforward application of the SPA is not possible.

@@

We introduce the theoretical framework of the ASCC, SPA,
and their combination, ASCC+SPA in Sec. \ref{4-1}. 
In Sec. \ref{4-2}, we provide some details in the application
of the ASCC+SPA to the multi-level pairing model.
We give the numerical results in Sec. \ref{4-3}, including
neutron pair vibrations in Pb isotopes.

@@

%\begin{figure}[tb]
% \begin{center}
%    \includegraphics[width=80mm, bb=0 0 550 180]{SPA.png}
% \end{center}
%  \caption{For a TDHFB trajectory, the correspondence between TDHFB phase space and collective subspace.}
%  \label{fig:correspondence}
%\end{figure}
%We try to combine SPA with ASCC. The key point is to find the correspondence between TDHFB phase space and collective subspace in a TDHFB trajectory (See Fig. \ref{fig:correspondence}). We consider the time-dependent state vector $\ket{Z(t)}$ is in a collective subspace. With pairing correlation, $\ket{Z(t)}$ can be expressed as

@@

\section{Result}
\label{4-3}
In the pairing model in Chap.~\ref{chap2},
the number of TDHFB degrees of freedom equals that of single-particle levels.
% including the constant of motion (pair rotation). We consider various systems, two-level system, three-level system, and Pb isotope system in each subsection respectively. Next, we discuss the dynamics in non-integrable (more than three-level) system.  
As there are two constants of motion, that is, the particle number and 
the energy, the system is integrable for one- and two-level models.
We first apply the ASCC+SPA method to an integrable two-level model,
then, to non-integrable multi-level models.

>>>>

Of course, single-particle motion picture cannot describe the nuclear system perfectly. To describe the nuclear structure more precisely, residual interactions need to be taken account. In 1950s, nuclear physicists found one of the most important residual interactions, pairing correlation. One of the clear evidences for pairing correlation detected from experiment is the separation energy difference between odd nuclei and even nuclei. Fig.~\ref{S_n} shows the separation energies for Sn, Sb, and Te isotopes as a function of neutron number. The isotopes in even number of neutron have about $2\sim3$ MeV larger separation energies than the isotopes in odd number of neutron.
It is odd-even staggering (OES) effect, which is detected in whole mass region of nuclei (Fig.~\ref{gap}). In this figure, $2\Delta$ corresponds to the separation energy difference between even-even nuclei and odd mass number nuclei. The approximate value is $2\Delta \approx 24A^{-1/2}$ MeV. It indicates a two-body attractive interaction should be included based on single-particle motion picture. In addition, if we examine the spin-parity of the ground states, we can find all of the even-even nuclei are $J^{\pi}=0^+$ states and most of the odd mass number nuclei have the same spin-parity as their single-particle orbit which the last nucleon fulfills in.
To explain these phenomena, it is natural to consider that the ground state of even-even nuclei consist of coupled $J^{\pi}=0^+$ pairs and the ground state of odd mass number nuclei consist of coupled $J^{\pi}=0^+$ pairs except one unpaired nucleon in last single-particle orbit. The two-body attractive interaction which couples two identical nucleons into $J^{\pi}=0^+$ state is called pairing correlation.

@@

First, we discuss the pairing rotation. In classical picture, pairing rotation is the motion which the system rotates in the U(1) gauge space (red bi-direction arrow in Fig.~\ref{pair_rot}). The motion is described by the coordinate, gauge angle $\Phi$, and the corresponding momentum, total particle number $N$. 
%It is similar to normal rotation in 3D space which described by the angles ($\theta_x, \theta_y, \theta_z$) and their conjugate angular momenta ($J_x, J_y, J_z$). 
Of course, $N$ is a constant of motion. In quantum picture, if we assume the ground state of $N_0$ particle system as the ``ground state'', the ground states in $N=N_0\pm2, N_0\pm4,\cdots$ correspond to the ``excited states'' in the pairing rotational mode. The excitation energy is $E_{\rm rot}^{\rm pair}=\frac{(N-N_0)^2}{2\mathcal{J}}$, where $\mathcal{J}$ is the pairing-rotational moment of inertia. In experiment, the pairing rotational band is expected to be observed in even-even isotope chain or even-even isotone chain in open-shell nuclei. A clear observation is found in Sn isotopes shown in Fig.~\ref{pair_rot}. Except a constant value and linear term with respect to mass number $A$, we can find the energies of ground states consist of a parabolic line - pairing rotational band. Recently, the systematic study for the pairing-rotational moment of inertia has been done in Ref. \cite{HN16}. The experimental moments of inertia obtained from two neutron(proton) separation energies have excellent correspondence with the theoretical moments of inertia obtained from NG mode in quasiparticle random phase approximation (QRPA) \cite{TV61}, not only for the spherical nuclei, but also for the deformed nuclei. Because the pairing-rotational moments of inertia are free from ambiguities caused from odd-even system, they are supposed to be excellent indicators of pairing correlation rather than pairing gaps.

@@

The signatures for pairing collective motion can also be observed from the two-neutron(proton) transfer reaction \cite{BHR73}, such as $(p, t)$, $(t, p)$, and $({}^3{\rm He}, n)$. The cross sections of two particle transfer reaction strongly depend on the matrix elements of $J^{\pi}=0^+$ pair-transition operators (e.g. pair-additional transition: $\sigma( P_{ad};0_{\alpha}^+(N) \to 0_{\beta}^+(N+2))\sim |\braket{N+2,\beta|\sum_k a^{\dag}_k a^{\dag}_{\bar{k}}|N,\alpha}|^2$). From theoretical prediction, the pairing correlation is known to enhance $\sigma(g.s.\to g.s.)$ and decrease $\sigma(g.s.\to 0_{ex}^+)$ significantly, hence, $\sigma(g.s.\to 0_{ex}^+)/\sigma(g.s.\to g.s.)\ll 1$ is expected in superfluid system. One example of $(p, t)$ reactions is demonstrated in Fig.~\ref{pair_rot}. As mentioned in previous paragraph, the ground states of even-even Sn isotopes consist of a pairing rotational band. In such situation, the intraband transition is dominant with respect to the interband transition. Therefore, the cross sections of two-neutron transfer reaction is consistent with theoretical prediction. In general, the ratio $\sigma(g.s.\to 0_{ex}^+)/\sigma(g.s.\to g.s.)$ for two-neutron(proton) transfer reaction are excellent indicators to describe the collectivity of pairing dynamics.

@@

\begin{figure}[tb]
 \begin{center}
    \includegraphics[width=100mm]{images/phonon.eps}
 \end{center}
 \begin{center}
    \includegraphics[width=100mm]{images/112Cd.eps}
 \end{center}
  \caption{$B(E2)$ values, relative to the $B(E2; 2_1^+\to 0_1^+)$ value, and energy spectra in phonon excitation picture. Upper panel: expected $B(E2)$ values for a quadrupole harmonic vibrator. Lower panel: data observed for ${}^{112}$Cd. Dashed transitions indicate an unobserved transition. Detailed information is discussed in Ref. \cite{GW10}.}
  \label{112Cd}
\end{figure}
The mystery of excited $0^+$ states exist not only in deformed nuclei, but also in spherical nuclei. A best example is the structure of $0_2^+$ or $0_3^+$ states in Cd isotopes observed in recent experiment \cite{GW10}. The lower part of Fig.~\ref{112Cd} shows the excitation energy and $B(E2)$ transition strength in ${}^{112}$Cd. Comparing with the typical values for the ideal quadrupole harmonic vibration in spherical nuclei shown in the upper part of Fig.~\ref{112Cd}, we can find that the observed excitation energy spectra have good coincidence with phonon excitation picture. The $0_2^+$ or $0_3^+$ states seem to be interpreted as two-phonon or three-phonon excitation. However, the $B(E2)$ values are quite different between the two, especially for the no observed decay from $0_{2ph}^+$ to $2_1^+$ state. Such phenomena are also seen in other Cd isotopes. The configuration of these $0^+$ states are supposed to be not pure phonon excitation, but proton pair-excitation between $2p_{1/2}$ and $1g_{9/2}$ orbits from the hint of the ground state configurations of Ag isotopes \cite{HW11}. The pairing should play an important role if the excited $0^+$ states are described by pair-excitation.

@@

%Large amplitude motion, Adiabatic approach
To describe the low-lying excited states, especially for the excited $0^+$ states, we need to introduce the large-amplitude motions beyond the linear regime.
Collective model, which is designed by a couple of collective coordinates and their collective potential and mass parameters, is useful for describing the large-amplitude dynamics. The most popular phenomenological collective model is Bohr collective model, five-dimensional collective Hamiltonian (5DCH) with the quadrupole deformation parameters $(\beta,\gamma)$ and the Euler angles $(\phi,\theta,\psi)$. It has been applied to describe the low-lying excited states in realistic nuclei, with the construction of the collective potential and mass parameters from the microscopic degrees of freedom \cite{NMMY16}. However, the pairing degrees of freedom are not explicitly included in the Bohr collective model. As discussed in Sec.~\ref{1-1-2}, the pairing vibrational degrees of freedom supposed to be a key ingredient to understand the low-lying  excited $0^+$ states. We have great interest for the collective coordinates describing the pairing corrective motions.

@@

\clearpage{\pagestyle{empty}\cleardoublepage}
\chapter{Requantization of TDHFB in integrable system}
\label{integrable}
As discussed in Sec.~\ref{constant}, there is only one time-dependent degree of freedom in two-level system. We know that one-dimensional system must be integrable system.
We obtained two achievements from the study for the integrable system. First, the requantization method, stationary-phase approximation to the path integral, can be applied straightforward to describe the collective excited $0^+$ states. We find that SPA is superior to the other requantization methods, especially for the calculation of two-particle transition strength. Second, the collective coordinates are explicit in integrable system. Using the explicit collective coordinates, we can examine the performance of the collective model which assumes the pairing gap parameter as the collective coordinate. The conclusion is the pairing gap parameter can only describe a part of collective space, and the applicability of the collective model is limited.

@@

We show an example of the classical trajectories in two-level system. Fig.~\ref{fig:phase_space} is the contour lines fulfilling (\ref{contour}), which correspond to the classical trajectories, with $\Omega_1=\Omega_2=\Omega=8$, $\nu=0$, and $N=2\Omega=16$ with various coupling constant $g$. The dimensionless coupling constant $x$ is defined as $x=2g\Omega/(\epsilon_2-\epsilon_1)$ (See Appendix \ref{two-level}). The phase transition from normal state to superfluid state occurs at $x=1$ in closed shell ($N=2\Omega$). At $x<1$ corresponding the left panel in Fig.~\ref{fig:phase_space}, the trajectory of ground state is $j=-4$ independent from $\phi$. It is consistent with the fully occupied lower level $n_1=N$ and the empty upper level $n_2=0$. For the dynamical states, the trajectories are open with gentle slopes. At $x>1$ corresponding the middle and right panels in Fig.~\ref{fig:phase_space}, the trajectory of ground state becomes a point at $\phi=0$, $j=j_0$ $(-4<j_0<4)$. With superfluid ground state, the dynamical states have two types, closed trajectory and open trajectory. The closed trajectory appears around $(\phi,j)=(0,j_0)$ at low-energy region. On the other hand, the open trajectory with steep slope appears at high-energy region relatively. The stronger pairing correlation, the higher excitation energy of the transition point from closed trajectory to open trajectory. The structure of the excited states between closed trajectory and open trajectory is quite different. They influence the matrix element of two-particle transfer reaction significantly. We will discuss the feature of the excited states in the following sections.

\subsection{TDHFB dynamics in adiabatic\footnote{In this subsection, the word ``adiabatic'' means that the collective velocity is small.} limit}
\label{ATDHFB}
In general collective motion, the frequency of the collective motion is much smaller than the frequency in single particle motion ($\hbar\omega\ll\epsilon_m-\epsilon_i$). The slow velocity of the collective motion enable us to expand the TDHFB Hamiltonian up to second order with respect to the collective momentum. We discuss the behavior of TDHFB dynamics in adiabatic approximation.\par
From (\ref{Hamiltonian2}), we find that the classical Hamiltonian is even function with respect to $\phi$. With time reversal $t\to-t$, (\ref{TDHFB_equation}) is invariant under $j\to j, \phi\to -\phi$, which indicates that $j$ is time-even and $\phi$ is time-odd. Thus, we switch the coordinate and momentum 
\begin{equation}
	(\phi,j)\to(j,-\phi) ,
	\label{switch}
\end{equation}
to make the coordinate time-even. The classical Hamiltonian under adiabatic approximation becomes
\begin{align}
  \mathcal{H}(j,\phi) &\approx V(j) + \frac{1}{2}B^{-1}(j)\phi^2,
  \label{ATDHFB_Hamiltonian}
\end{align}
where 
\begin{align}
	V(j) = \mathcal{H}(\phi=0,j) 
	=& \sum_{\alpha=1,2} \epsilon_{\alpha}\nu_{\alpha} + \sum_{\alpha=1,2} 2\epsilon_{\alpha}j_{\alpha} - g\sum_{\alpha=1,2} \left( (\Omega_{\alpha}-\nu_{\alpha}) j_{\alpha} - j_{\alpha}^2 +\frac{j_{\alpha}^2}{\Omega_{\alpha}-\nu_{\alpha}} \right) \nonumber 
	\label{potential_two_level} \\
	&- 2g\sqrt{j_1j_2(\Omega_{1}-\nu_{1}-j_{1})(\Omega_{2}-\nu_{2}-j_{2})} \\
	B^{-1}(j) = \left. \frac{\partial^2\mathcal{H}}{\partial\phi^2} \right|_{\phi=0}
	=& 2g\sqrt{j_1j_2(\Omega_{1}-\nu_{1}-j_{1})(\Omega_{2}-\nu_{2}-j_{2})} .
	\label{mass_two_level}
\end{align}
It is obvious that $V(j)$ is potential and $B(j)$ is mass parameter. We define a classical trajectory obtained from (\ref{ATDHFB_Hamiltonian}) as adiabatic TDHFB (ATDHFB) trajectory. Because the closed trajectories appear around the energy minimum point $(\phi,j)=(0,j_0)$, we expect that adiabatic approximation performs excellent for closed trajectories.
\begin{figure}[tb]
 \begin{center}
  \includegraphics[width=60mm,angle=-90]{images/phase_compare.eps}
 \end{center}
 \caption{Energy contour plot in the same system as Fig.~\ref{fig:phase_space}. Black solid lines correspond to TDHFB trajectories and gray dashed lines correspond to ATDHFB trajectories. 
	}
 \label{fig:phase_compare}
\end{figure}

We compare the difference of the classical trajectories between TDHFB and ATDHFB. Fig.~\ref{fig:phase_compare} shows both TDHFB and ATDHFB trajectories in the same system as in Fig.~\ref{fig:phase_space}. For weak pairing correlation ($x<1$), both results are similar, especially at the low-energy region. For strong pairing correlation ($x>1$), both results are almost identical for closed trajectories. The opened trajectories are also well reproduced near the closed trajectories. The deviation becomes large at the high-energy region. In conclusion, ATDHFB is excellent approximation at low-energy region for both open and closed trajectories.

\section{Requantization methods}
\label{3-2}
To obtain quantum states from TDHFB dynamics, quantization of the TDHFB trajectories, namely, requantization is necessary procedure. We emphasize that there are infinite quantum pictures corresponding to one classical picture. Different quantization methods lead to different quantum states. The straightforward requantization method is canonical quantization, which is commonly applied into nuclear collective model. Here, we apply the canonical quantization to describe the excited states. However, we find that canonical quantization has serious problems to describe the pairing dynamics. Such problems are discussed in Sec.~\ref{3-3}. To solve or avoid the problems, alternative requantization methods, Fourier decomposition and stationary-phase approximation to the path integral, are also introduced. In this section, we explain each requantization method in detail, respectively.

@@

Let us note the periodic conditions
of the coordinates and the quantization condition.
We have three types of periodic conditions. (1) $\Phi\rightarrow \Phi\pm 2\pi$ with fixed $\phi$. The quantization condition (\ref{EBK}) becomes
\begin{equation}
	2\pi J = 2k \pi ,
\end{equation}
which determines nothing but the total particle number $J$. (2) $\phi\rightarrow \phi$ with fixed $\Phi$. The quantization condition (\ref{EBK}) becomes
\begin{align}
	\oint j d\phi = 2k \pi ,
\end{align}
which determines the excited states for closed TDHFB trajectories (e.g., 
\ref{fig:phase_space}). (3) $(\Phi,\phi)\rightarrow (\Phi\pm \pi,\phi\pm 2\pi)$ and
$(\Phi,\phi)\rightarrow (\Phi\pm \pi,\phi\mp 2\pi)$.
because the two original variables, $(\chi_1,\chi_2)$,
are independent periodic variables of the period $2\pi$,
in addition to the trivial periodicity of $2\pi$ for $\Phi$ and of $4\pi$
for $\phi$.
The former (latter) corresponds to $\chi_2\pm 2\pi$ ($\chi_1\pm 2\pi$)
with $\chi_1$ ($\chi_2$) being fixed.
The quantization condition (\ref{EBK}) becomes
\begin{align}
 \mathcal{T}_{J,E_k}(\Phi\pm\pi,\phi\pm 2\pi;J) &= \mathcal{T}_{J,E_k}(\Phi,\phi;J) + 2m\pi \nonumber \\
	 \Leftrightarrow \quad\quad \pm J \pi + \int_{-\pi}^\pi j d\phi
	&= 2m\pi ,
\end{align}
which determines the excited states for open TDHFB trajectories (e.g., Fig.~\ref{fig:phase_space}).
This leads to the following:
\begin{equation}
	\int_{-\pi}^\pi j d\phi = 
	\begin{cases}
		2k\pi & \mbox{ for } J=2n ,\\
		(2k+1)\pi & \mbox{ for } J=2n+1 ,
	\end{cases}
	\label{EBK_2}
\end{equation}
where $m$, $k$ and $n>0$ are integer numbers.

@@

From the study of various requantization methods for the two-level pairing model,
the stationary-phase approximation (SPA) to the path integral \cite{SM88} 
gives quantitative results not only for the excitation energies, 
but also for the wave functions and two-particle-transfer strengths.
The quantized states obtained in the SPA have two advantages;
first, the wave functions are given directly in terms of
the microscopic degrees of freedom, and second,
the restoration of the broken symmetries are automatic.
In the pairing model, the quantized states are eigenstates of the particle-number
operator.
On the other hand, applications of the SPA have been limited to
integrable systems.
This is because we need to find separable periodic trajectories on a classical
torus.
Since the nuclear systems, of course, correspond to non-integrable systems,
a straightforward application of the SPA is not possible.

@@

We introduce the theoretical framework of the ASCC, SPA,
and their combination, ASCC+SPA in Sec.~\ref{4-1}. 
In Sec.~\ref{4-2}, we provide some details in the application
of the ASCC+SPA to the multi-level pairing model.
We give the numerical results in Sec.~\ref{4-3}, including
neutron pair vibrations in Pb isotopes.

@@

%\begin{figure}[tb]
% \begin{center}
%    \includegraphics[width=80mm, bb=0 0 550 180]{SPA.png}
% \end{center}
%  \caption{For a TDHFB trajectory, the correspondence between TDHFB phase space and collective subspace.}
%  \label{fig:correspondence}
%\end{figure}
%We try to combine SPA with ASCC. The key point is to find the correspondence between TDHFB phase space and collective subspace in a TDHFB trajectory (See Fig.~\ref{fig:correspondence}). We consider the time-dependent state vector $\ket{Z(t)}$ is in a collective subspace. With pairing correlation, $\ket{Z(t)}$ can be expressed as

@@

\section{Result}
\label{4-3}
We first apply the ASCC+SPA method to an integrable two-level model,
then, to non-integrable multi-level models.
